quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-13 paper-quollr.Rmd
Initial update for link plot section
Added show_link_plots description

There are mainly two interactive link plots can be generated. `show_link_plots()` helps to examine the fit. The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, and high-dimensional model data, and edge data where the from and to links of the edges.

>>>>

There are mainly two interactive link plots can be generated.

To support interactive evaluation of how well the $p\text{-}D$ model captures the structure of the high-dimensional data, we introduce `show_link_plots()`. This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space. Both views are interactively linked, enabling users to explore.

Before visualization, the input data must be prepared using the `comb_all_data_model()` function. This function combines the high-dimensional data (`highd_data`), the NLDR output (`nldr_data`), and the bin-averaged high-dimensional model representation (`model_highd`) aligned to the 2D bin layout (`model_2d`):

This combined dataset includes both the original observations and the bin-level model averages, labeled with a `type` variable for distinguishing between them.

The function `show_link_plots()` generates two side-by-side, interactively linked plots; a 2D scatterplot of the NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the `langevitour` package), displaying both the data and the model.

These two views are linked using `crosstalk`, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection. This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.

### 2025-05-13 paper-quollr.tex
Initial update for link plot section
Added show_link_plots description

There are mainly two interactive link plots can be generated. \texttt{show\_link\_plots()} helps to examine the fit. The function requires several arguments: points data which contain Non-linear dimension reduction data, high-dimensional data, and high-dimensional model data, and edge data where the from and to links of the edges.

>>>>

There are mainly two interactive link plots can be generated.

To support interactive evaluation of how well the \(p\text{-}D\) model captures the structure of the high-dimensional data, we introduce \texttt{show\_link\_plots()}. This visualization combines two complementary views: the nonlinear dimension reduction (NLDR) representation and a dynamic tour of the model ovelaid the data in the high-dimensional space. Both views are interactively linked, enabling users to explore.

Before visualization, the input data must be prepared using the \texttt{comb\_all\_data\_model()} function. This function combines the high-dimensional data (\texttt{highd\_data}), the NLDR output (\texttt{nldr\_data}), and the bin-averaged high-dimensional model representation (\texttt{model\_highd}) aligned to the 2D bin layout (\texttt{model\_2d}):

This combined dataset includes both the original observations and the bin-level model averages, labeled with a \texttt{type} variable for distinguishing between them.

The function \texttt{show\_link\_plots()} generates two side-by-side, interactively linked plots; a 2D scatterplot of the NLDR representation, and a dynamic projection tour in the original high-dimensional space (using the \texttt{langevitour} package), displaying both the data and the model.

These two views are linked using \texttt{crosstalk}, allowing interactive selection of points in the NLDR plot to highlight corresponding structures in the high-dimensional projection. This setup facilitates the diagnosis of local distortion, structural artifacts, and model fit quality.
